Eco-evolutionary dynamics of finite populations from first principles

### 2023-02-21 ind_files/06_unification_discussion.tex
Correct minor typo in discussion of variance Price for neutral evol

The $\overline{\tau}\sigma^2_{f}$ term represents a loss of diversity due to stochastic extinctions (i.e. demographic stochasticity).To see this, let us consider the case in which this is the only force at play. Let us imagine we have a population of asexual organisms and each $f_i$ is simply a label or mark arbitrarily assigned to individuals in the population at the start of an experiment/observational study and subsequently passed on to offspring - For example, a neutral genetic tag in a part of the genome that experiences a negligible mutation rate. Let us set $\mu = 0$ so that the labels cannot change between parents to offspring. This means that we have $M_{\sigma^2_f}(\mathbf{p},N_K) \equiv 0$. Further, since the labels are arbitrary and have no effect whatsoever on the biology of the organisms, we have $\textrm{Cov}\left(w,(f - \overline{f})^2\right) \equiv \textrm{Cov}\left(\tau,(f - \overline{f})^2\right) \equiv 0$. Since the labels do not change over time, we also have $\textrm{Cov}\left(\partial f/\partial t,f\right) = 0$. From \eqref{stochastic_Price_variance}, we see that in this case, the variance changes as

@@

We were able to decompose the expectation on the RHS because the label $f$ is completely arbitrary and thus independent of both $\overline{\tau}$ and $N_K(t)$. Equation \eqref{neutral_example_for_variances} is a simple linear ODE and has the solution

@@

\mathbb{E}[\sigma^2_f(t)] = \sigma^2_f(0)\exp\left(-\mathbb{E}\left[\frac{\overline{\tau}}{KN_K}\right]\right)

@@

which tells us that the population is expected to lose its diversity of labels at an exponential rate. Populations with higher mean turnover $\overline{\tau}$  and/or lower population size $KN_K$ lose diversity faster, which should be intuitive since these populations experience more extreme stochastic events and are therefore more likely to experience stochastic loss of labels. Note that \emph{which} labels/individuals are lost is entirely random (since all labels are arbitrary), but nevertheless, labels tend to be stochastically lost until only a small number of labels remain in the population.

>>>>

The $\overline{\tau}\sigma^2_{f}$ term represents a loss of diversity due to stochastic extinctions (i.e. demographic stochasticity). To see this, it is instructive to consider the case in which this is the only force at play. Let us imagine a population of asexual organisms in which each $f_i$ is simply a label or mark arbitrarily assigned to individuals in the population at the start of an experiment/observational study and subsequently passed on to offspring - For example, a neutral genetic tag in a part of the genome that experiences a negligible mutation rate. Let us set $\mu = 0$ so that the labels cannot change between parents to offspring. This means that we have $M_{\sigma^2_f}(\mathbf{p},N_K) \equiv 0$. Further, since the labels are arbitrary and have no effect whatsoever on the biology of the organisms, we have $\textrm{Cov}\left(w,(f - \overline{f})^2\right) \equiv \textrm{Cov}\left(\tau,(f - \overline{f})^2\right) \equiv 0$. Since the labels do not change over time, we also have $\textrm{Cov}\left(\partial f/\partial t,f\right) = 0$. From \eqref{stochastic_Price_variance}, we see that in this case, the variance changes as

@@

where we have decomposed the expectation of the product on the RHS into a product of expectations, which is admissible since the label $f$ is completely arbitrary and thus independent of both $\overline{\tau}$ and $N_K(t)$. Equation \eqref{neutral_example_for_variances} is a simple linear ODE and has the solution

@@

\mathbb{E}[\sigma^2_f(t)] = \sigma^2_f(0)\exp\left(-\mathbb{E}\left[\frac{\overline{\tau}}{KN_K}\right]t\right)

@@

which tells us that the population is expected to experience an exponential loss in the diversity of labels over time. The rate of loss is $\mathbb{E}\left[\overline{\tau}(KN_K)^{-1}\right]$, and thus, populations with higher mean turnover $\overline{\tau}$  and/or lower population size $KN_K$ lose diversity faster. This is because such populations experience more extreme stochastic events and are therefore more likely to experience stochastic loss of labels. Note that \emph{which} labels/individuals are lost is entirely random (since all labels are arbitrary), but nevertheless, labels tend to be stochastically lost until only a small number of labels remain in the population.
